# JOUR405: Statistics for Journalists
## Midterm Exam - Spring 2025

Name: Ryan Martin

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences. Start by loading the tidyverse and any other libraries you think you might need.

```{r}
library(tidyverse)
```


## Part 1: Restaurant Health Inspections (15 points)

You want to understand how restaurants in Montgomery County are performing on health inspections. The first dataset contains restaurant health inspection scores for restaurants in Montgomery County. The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv` and complete these tasks:
```{r}
Health_Inspections <-read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```


### Tasks:
1. Calculate the mean and standard deviation of compliance scores (5 points)
2. Create a histogram of the compliance scores with a vertical line showing the mean (5 points)
3. Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? (5 points).

```{r}
Health_Inspections |> summarize(mean = mean(compliance_score), sd = sd(compliance_score))
Health_Inspections |> 
  ggplot() +
  geom_histogram(aes(x = compliance_score), binwidth = 3) +
  geom_vline(aes(xintercept =mean(compliance_score)), color ='red', linetype = "dashed")

```
When looking at the compliance scores of health inspect reports both through the mean, standard deviation, and created histogram, we can see that establishments scored very highly with consistency on the reports. The high mean with a lower standard deviation shows the rarity of which it was that establishments did not score highly, data that's reflected in the histogram that shows a clear trend in which establishments nearly always earned a score in the high 90s.

## Part 2: High School Athletics (25 points)

You are reporting a story about high school sports participation in Maryland and want to see if there are differences between boys and girls. The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and sex. Load the data from: `https://raw.githubusercontent.com/example/md_hs_sports_2024.csv` and complete these tasks:

```{r}
library(tidyverse)
```
```{r}
Athletics <-read_csv ("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```



### Tasks:
1. Calculate the correlation between boys' and girls' participation (5 points)
2. Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. (5 points)
3. Create a scatterplot showing this relationship, adding a line of best fit (5 points)
4. In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

```{r}
Athletics |> summarize(correlation = cor(boys, girls, method="pearson"))
Athletics <-Athletics |> mutate(total = boys + girls)
Athletics <-Athletics |> mutate(girls_pct = girls / total)
Athletics |> 
  ggplot()+
  geom_smooth(aes(x=total, y=girls_pct), method= "lm") +
  geom_point(aes(x=total, y=girls_pct))

```
Both the correlation and scatter plot with line of best fit shows that there's equity in Maryland high school sports participation but that equity is dependent on the number of total athletic participants and, in turn, likely dependent on the total number of students per school. The correlation numbers show that an increase in the athletic participation of female and male students is highly correlated, data that's reinforced by the scatter plot that expands to show how female participation slightly decreases as the total number of athletic participants increases. School districts below the line need to be more equitable with their athletic resources and the districts worth examining further include the outliers across the X intercept (showing huge total participation with an average girl participation) and Y intercept (showing huge girl participation among an average total population.) to see what they're doing differently than other districts.

## Part 3: Public Transit Ridership (20 points)

You are investigating public transit ridership in the Washington, D.C. area and want to understand the patterns of daily bus and rail ridership. The third dataset contains daily bus and rail ridership totals from WMATA for the past year. Load the data from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv and do the following:

```{r}
Transportation <-read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```


### Tasks:
1. Calculate the average bus and rail ridership and standard deviation using summarize() (5 points)
2. Using the process you used in class, take a random sample daily ridership numbers and calculate the sample means and deviations for bus and rail. The number in the sample is up to you, but explain why you chose what you did. Compare this to the stats you generated in step 1. (5 points)
3. Using group_by() and summarize(), calculate the means for bus and rail ridership for each weekday. Describe the overall pattern of ridership for bus and rail - which days stand out and why? Are there differences between bus and rail in the standard deviation values? (10 points)


```{r}
Transportation |> summarize(sd = sd(bus))
Transportation |> summarize(sd = sd(rail))

set.seed(1234) # Setting seed for reproducibility
  sample42 <- Transportation |> 
      sample_n(42)
  
  set.seed(1234) # Setting seed for reproducibility
  sample128 <- Transportation |> 
      sample_n(128)
  
sample42 |> summarize(mean = mean(bus), sd = sd(bus))
sample42 |> summarize(mean = mean(rail), sd = sd(rail))

sample128 |> summarize(mean = mean(bus), sd = sd(bus))
sample128 |> summarize(mean = mean(rail), sd = sd(rail))


Weekdays_bus <- Transportation |> 
    group_by(weekday) |>
    summarize(averages_bus =mean(bus))

Weekdays_rail <- Transportation |> 
    group_by(weekday) |>
    summarize(averages_rail =mean(rail))

```
I chose to do a sample of 42 observations and 128 observations because I was interested in seeing what roughly 10% of the overall data (representative with 42 observations) and roughly 30% of the overall data (representative with 128 observations) would demonstrate. Looking at the stats, we can see that the higher of a random sampling done, the closer to the accurate mean and standard deviation we get is.

We can look at the overall pattern for both rail and bus ridership and notice that, generally, Tuesdays through Thursdays tend to be the busiest days for both bus and rail ridership before the weekend sees less patrons using this transportation. The weekend days stick out because of their very low patronage compared to the week; the bus has a much higher standard deviation than the rail.

## Part 4: Maryland Car Theft Rates (20 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

```{r}
MD_Car_Theft <-read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```


### Tasks:
1. Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate (5 points)
2. Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median, and what percentage of all car thefts occur in those counties? (5 points)
3. Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? (10 points)


```{r}
MD_Car_Theft <-MD_Car_Theft |> mutate (Rate = `2023` / population)

MD_Car_Theft |> mutate(Theft_Median = median(Rate))


```
Anne Arundel, Baltimore City, Baltimore, Howard, Montgomery, and Prince George's are counties that are all above the median for car theft in Maryland with a 0.0018276091, 0.0123667473, 0.0028566855, 0.0020873513, 0.0021908384, 0.0089432222 rate. Close to 60% of all of Maryland's car theft crimes occur in six counties, respectively. 

With these calculations in mind, we have a clearer understanding about the distributions of car thefts in Maryland and we can see how concentrated these kinds of theft are within these three counties. The lede to our story would be that even though there are close to 25 counties in Maryland, over haldf of the car theft related crimes comes from just these six counties, or less than a fourth of the total Maryland counties.

## Part 5: Data Analysis Scenario (20 points)

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences (no code!) explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)

I would look at the mean and standard deviations in response times for police, fire and ambulance calls monthly over a 12 month span to verify if these claims of a worsening response times are true or not.

2. What visualizations would help readers understand the trends? (5 points)

This data could be well visualized using a scatter plot with a line of best fit to show if there is or is not a growing trend for this claim.

3. What additional context or data would you need to make this a complete story? (5 points)

To complete this story, additional context witjin the police, fire, and ambulance departments would be necessary — background on if these departments have had to lay off personnel before, if there's been a decrease in the quality/efficiency of communication between emergency response dispatchers and these departments compared to last year, etc.


### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Before submitting, clear your environment and run the entire notebook

Good luck!
