---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```

### Task 2: Load 1,000 rows of data with the ages of people from 18-80
```{r}
set.seed(42) # Setting seed for reproducibility
population  <-tibble(
  age = sample(18:80, 1000, replace = TRUE)
)

# View the first few rows
head(population)
```


### Task 3: Sample 100 rows of data from the dataset
```{r}
sample_100 <- population |> 
  sample_n(100)

sample_50 <- population |> 
  sample_n(50)

sample_10 <- population |> 
  sample_n(10)

```



### Task 4: Calculate the mean age and standard deviation of the sample
```{r}
sample_100 |> summarize(mean = mean(age), sd = sd(age))

sample_50 |> summarize(mean = mean(age), sd = sd(age))

sample_10 |> summarize(mean = mean(age), sd = sd(age))
```


### Task 5: Calculate the mean age and standard deviation of the population
```{r}
population |> summarize(mean = mean(age), sd = sd(age))
```


### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

They get less and less similar while getting less and less accurate as the sample size goes down. When the sample size goes down, the results differ more from the true mean. It wasn't that far off, though, showing we can understand public opinion in facets easier than we previously thought if the methodology includes the diversity represented in the entire population.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
