---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse

```{r}
library("tidyverse")
```





### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv

```{r}
umd_fees <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```


### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (`Fall.2024`- `Fall.2021`)/`Fall.2021` * 100)
```


### Task 4: Describe our results

Write a few sentences about what you found in the data. What's the story here?


When calculating the percent change in each category, it's really interesting to see the categories that vastly changed over the years. Here, I think the story is that every category increased in price as the overall fees went up annually with the exception of Athletics. Athletics was always consistent, and until recently, was the highest fee that students paid. For this, I think the story is the high yet consistent price that ducks just below $400 every year for athletics.


When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
