---
title: "Final_Project" - Ryan Martin
output: html_document
---

```{r}
library(tidyverse)

#Load Dataset

Transfer_Stats <-read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTxh6lyUk5C5u2dKVpm9DgqXXlOiGai6rJWxNicrOPw9z_OBq-3b8_T2qkvaEEnBNmvTPA8VKm5zDUa/pub?gid=69510539&single=true&output=csv")

Non_Transfer_Stats <-read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTxh6lyUk5C5u2dKVpm9DgqXXlOiGai6rJWxNicrOPw9z_OBq-3b8_T2qkvaEEnBNmvTPA8VKm5zDUa/pub?gid=2050881943&single=true&output=csv")
```

# Task 1: Explanation of Final Project

After conversations with Prof. Willis, I have adjusted my final project data and overarching question. When reflecting to former Maryland men's basketball coach Kevin Willard's tenure in College Park, a consistent theme with his teams was the need for improved and increased thrre-point shooting. So, to Willard's credit, Maryland attacked the transfer portal and often brought in players who were known as shooting three-point shots well at their previous schools. But when these players got to Maryland, often from smaller schools that played lower levels of competition, how successful were they at continuing their three-point efficiency going up against Big Ten Competition. This project evaluates all of that, looking closer into the twelve transfers Willard brought in during his time here and seeing how their three-point shooting at their previous school compared to their three-point shooting at Maryland. Collectively, the players actually shot better at Maryland than they combined to shoot at their previous institutions — so why is that?

# Task 2: Summarizing Transfer Data
```{r}
Transfer_Stats|> summarize(mean_change = mean(Change),
                         median_change = median(Change),
                         min_change = min(Change),
                         max_change = max(Change))
```
Firstly, after data collection, we can summarize the data for Maryland men's basketball's transfers relevant to the qualified players' change in three-point percentage between the final season at their old institution and their succeeding career three-point percentage once getting to the Big Ten at Maryland. Above is a simple data summary that, for the 10 qualified Maryland men's basketball transfers (qualified meaning they attempted 10 or more three-point shots during their time with Maryland) that were added to Maryland men's basketball's roster when Kevin Willard was the head coach, shows the mean, median, minimum, and maximum data for the percentage change as explained above. This data shows that, for consistent three-point shooters Willard brought over during his time at Maryland, the trend for changes in three-point shooting when players transferred over was shooting very, very slighly worse. A -0.0482 mean change and -0.0175 median change shows that the average change in a qualified players' shooting percentage is a 4.82% drop at Maryland compared to the previous year at their past university. A -1.75% median change shows that outliers are present in the data, but reinforces how the change in an average qualified players' shooting percentage is minimal after transferring to Maryland. This data must come with the aestrisk that there are so many other factors to consider for why a player's three-point shooting percentage would change once they get to Maryland: more/less playing time, more/less three-pointers attempted, etc. But this data is useful in seeing how the average qualified player's three-point shooting percentage gets adjusted after transferring to Maryland.

# Task 3: Developing Scatter Plots to compare Transfer to Non-Transfer Data 

```{r}
Transfer_Stats |>
  ggplot(aes(x = `Percentage.After`, y = `Change`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Three-Point Percentage After Transferring", y = "Change in Three-Point Percentage after Transferring", title = "Three-Point Shooting Changes After Transferring In, Maryland Men's Basketball")

Non_Transfer_Stats |>
  ggplot(aes(x = `Percentage`, y = `Made.over.Attempted`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Three-Point Percentage At Maryland", y = "Three-Pointers Made Over Attempted at Maryland", title = "Three-Point Shooting Trends, Maryland Men's Basketball Non-Transfers")
```
Now, with the context of the data summary to give us a bit of a fuller picture into the average qualified Maryland men's basketball transfers' three-point shooting numbers at his old school compared to Maryland, we can utilize scatter plots as a visualization to compare data for transfer players and three-point shooting data for qualified players (again, 10+ three-pointers attempted at Maryland) who never transferred in during Willard's time at Maryland. These scatter plots don't have the same Y-axis — the transfer data here shows the change in three-point percentage after transferring vs. overall three-point percentages after transferring, while the non-transfer data shows a less coordiated visual of three-pointers made/attemped over the percentage number for that shooting-percentage. The matching x-axis is the most important variable to compare across the two visualizations. Here, you can see that the 10 qualified transfer players shot consistently higher percentages (representative with more samples recording higher values on the x-axis) than the 11 qualified non-transfer players did during the span of Kevin Willard's tenure at Maryland. This represents the first time that we've compared the transfer to non-transfer data, a key piece of information that can help a story on if Willard was successful or not at utilizing the transfer portal vs. traditional recruiting to bring in Maryland men's basketball's most efficienct three-point shooters.

# Task 4: Showing the Data in a Model

```{r}
Transfer_Data_Model <- lm(`Percentage.After` ~ `Change`, data = Transfer_Stats)

summary(Transfer_Data_Model)

```
Next, it could be useful to turn this transfer data for qualified men's basketball players into a predictive model that gives more descriptive stats and helps to see two things. One, if future players that Willard might have brought in with Maryland men's basketball could have in fact continued the trend of qualified players shooting slightly worse at Maryland compared to their other school. And two, seeing how singularly the correlation is between the players' shooting percentage after getting to Maryland being worse just because of the transfer. Here, the model with a high p-value says we as data consumers need to take into account the other factors we mentioned are at play for why a player's shooting percentage would differ at Maryland compared to at his previous institution. The change variable used shows a high degree of probable error in the data, showing how the data reported leaves too many loose ends untied to truly report on storylines here for Maryland men's basketball.

# Task 5: Making a Paired T-Test to further scatter plots
```{r}
paired_test <- t.test(Transfer_Stats$Change, Transfer_Stats$Percentage.After, paired = TRUE)

print(paired_test)
```
To progress off of the scatter plots for the qualified Maryland men's basketball transfers from earlier, the best and most telling test for this kind of data would be a paired T-test for the transfers to see if there's a correlation between them changing schools and them shooting either better or worse three-point percentages. A paired t-test will take the congregate mean of the players' mean three-point percentage at their previous schools and see if there's a statistically significant difference with that and their new three-point percentage at Maryland. Here, the paired t-test uncovers that the data is too spread out to be conclusive over the correlation. There shows to be a trend that qualified players shoot worse at Maryland compared to their old universities, but a high p-value and wide range at the 95% confidence interval show data inefficiencies are represented through the paired t-test.

# Task 6: Drafting A Story Pitch based on Final Analysis
With the tests and reported data considered, we as journalists have a foundation for a story here on qualified Maryland men's basketball transfers shooting significantly better at Maryland than qualified players who never transferred in during the Kevin Willard era, yet these transferring-in players showed worse three-point shooting percentages at Maryland compared to at their previous institution. My analysis proves this was best shown early in the notebook, specifically at Task 2 when summarizing the transfer data and in Task 3 when I visually looked at qualified Maryland men's basketball transfers three-point shooting percentage at Maryland compared to that of qualified Maryland men's basketball players who never transferred in under Willard. These were the main findings of the report and prove the original hypothesis and claim from Willard himself that, during his time as head coach of Maryland, he strategically brought in players who were known as three-point shooters at their previous institutions, welcoming these players into Maryland in hopes of improving the Terps' overall team three-point shooting. And it worked, qualified players who transferred in show to have shot more efficiently than the players who were recruited traditionally as high schoolers when Willard was here. But there are too many other factors at play, a facet of the data that was exposed in Task 5 during the Paired T-Tests and in Task 4 when the transfer data was modeled. The data becomes overly simplified, too, when comparing unequal and different variables between the "Transfer_Stats" data and the "Non_Transfer_Stats" data. More symmetry was needed across the two data sets to better understand if the story is how much better the transfers shoot at Maryland than non-transfers because they are simply better shooters, or if the resources, coaching, etc. at Maryland offers that ecosystem where transfers will shoot more efficiently than non-transfers. I think that's where the future reporting comes in after this data. There are many ideas that Prof. Willis and I floated out there when first thinking of this data; I'd like to take this information given and report to my editor an interest/story pitch in comparing more common variables across transfer and non-transfer players. How did these players shoot at home vs. on the road before they got to Maryland and then once they got to Maryland? Were those splits consistent with the same splits for non-transfer players who have spent their whole careers at Maryland? Or, if we can again back that transfer players shot more efficiently than non-transfer players did under Willard at Maryland, the future for reporting here can go ask "Why?" What was different about the shooting backgrounds for these transfers vs. non-transfer players for the transfers to be better shooters? Was their value for the transfers in stopping at a previous collegiate program to work on their shooting at the D1 level before making the move (usually) up into the Big Ten at Maryland? There are plenty of ways a data journalist, editor, and publication could develop this data into seeing a fuller picture into Willard's emphasis on three-point shooting from the transfer portal when he was the head men's basketball coach at the University of Maryland.
