---
title: "HW9_TestScores"
name: Ryan Martin
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Did a New Reading Program Lead to Better Scores?

The superintendent recently claimed that a new reading program has improved third-grade reading scores across the school district.

Before the program, third-grade students in the district averaged 72.6 points on standardized reading tests with a standard deviation of 4.8 points.

After implementing the program for one semester, you collected scores from 12 randomly selected classrooms:
74, 76, 73, 75, 78, 77, 74, 79, 75, 76, 77, 75

As a journalist, you need to determine: **Is there statistical evidence that reading scores have actually improved?**

## Task 1: Organize your data and initial assessment

Before you can run this codeblock, you will need to fill in a value where it says REPLACE_ME. That value can be found in the introduction.

```{r}
# Known information about reading scores before the new program
prior_mean <- 72.6 # average score
prior_sd <- 4.8     # standard deviation

# Reading scores after implementing the new program (12 classrooms)
new_scores <- c(74, 76, 73, 75, 78, 77, 74, 79, 75, 76, 77, 75) # Replace with the actual scores

# Create a journalist-friendly dataset
score_data <- tibble(
  classroom = paste("Classroom", 1:12),
  reading_score = new_scores
)

# View the data
score_data
```

### Reflection Question 1:
Based on just looking at the score_data dataframe, have test scores improved? How can you tell?

In general, you can see that test scores have improved because the second half of classes have a mean thats higher than the prior mean.


## Task 2: Calculate key statistics

Like Task 1, you will need to replace values where it says REPLACE_ME before running any code.


```{r}
# Calculate statistics based on the new reading scores
new_stats <- score_data |> 
  summarise(
    mean = mean(72.6),
    sd = sd(4.8),
    n = n()
  )

new_stats
```

### Reflection Question 2:
Looking at the mean and standard deviation of the new scores compared to the previous statistics, what initial observations can you make? What questions might these statistics raise for your reporting?

The new mean and standard deviation show a progression in the average school test scores, yet it doesn't show the reasons why or what could have led to that, leasding to more questions for my reporting.

## Task 3: Create a column chart

As before, replace any values marked REPLACE_ME based the instructions.


```{r}
# STUDENT TASK: Choose an appropriate fill color for the bars
my_fill_color <- "royalblue" # Replace with a color name like "royalblue", "darkgreen", etc.

# Create a visualization comparing new scores to the previous average
score_data |> 
ggplot(aes(x = classroom, y = reading_score)) +
  geom_col(fill = my_fill_color, alpha = 0.8) +
  geom_hline(yintercept = prior_mean, color = "darkred", size = 1, linetype = "dashed") +
  annotate("text", x = 2, y = prior_mean - 1, 
           label = "Previous Average (72.6)", hjust = 0, fontface = "bold", color = "darkred") +
  labs(
    title = "Reading Scores After New Program Implementation",
    subtitle = "Horizontal line shows previous district average of 72.6 points",
    x = NULL,
    y = "Reading Test Score",
    caption = "Source: District Assessment Data"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### Reflection Question 3:
Examine the chart you created, and suggest a better title based on the results of the data, not a description.

Based on the chart we made and the resulting data, a better title here would be that the new program being implemented has led to inconclusive results on if they've improved reading scores.

## Task 4: Perform a hypothesis test

This is where we formally test the superintendent's claim that reading scores have improved. Fill in the REPLACE_ME values as needed, beginning with your hypotheses.

**Hypotheses:**
Null: The new reading program being implemented does not improve test scores.
Alternative: The new reading program being implemented improves test scores.

```{r}
# Set the significance level for your test
alpha_level <- .05 # Replace with the appropriate value

# Perform a one-sample t-test
# Since we want to know if scores improved (increased), we use a one-sided test (alternative = "greater")
t_test_result <- t.test(
  score_data$reading_score,
  mu = prior_mean,
  alternative = "greater"
)

# Display the results
t_test_result
```

### Reflection Question 4:
What does the p-value tell you, and what doesn't it tell you? How would you explain these results to a non-technical audience while maintaining accuracy?

Here, the p-value tells us that the perceived result of improved test scores is not well correlated with the new implementation of a reading program. It doesn't tell us what other factors would lead to improved test scores if not for the reading program. You could explain this simply to an audience by maintaining accuracy by generalizing your findings, saying how the data is inconclusive because the high p-value shows there's inconsistencies with a clear trend. Also, because the old average is identical to the new one, we see there's practically no change.

## Task 5: Interpreting the results for your news story

Let's gather all of the important stats we'll need in one place, so we can look at the prior average, the new scores and the results of the t.test, including the confidence interval. Replace any values where it says REPLACE_ME.


```{r}
# Get the p-value
p_value <- t_test_result$p.value

# Calculate the 95% confidence interval
ci <- t.test(score_data$reading_score)$conf.int

# Create a tibble to display the key statistics for your story
story_stats <- tibble(
  `Previous average` = prior_mean,
  `New average` = mean(72.6),
  `Improvement` = mean(new_scores) - prior_mean,
  `Percent change` = round(((mean(new_scores) - prior_mean) / prior_mean) * 100, 1),
  `p-value` = p_value,
  `Lower bound` = ci[1],
  `Upper bound` = ci[2],
  `Confidence level` = "95%"
)

# Display the key statistics
story_stats
```

## Conclusion

### Reflection Question 5:
Based on these statistics, what would be your headline and lead paragraph for this story? Is there evidence to support the superintendent's claim?

My headline would be that "New Reading Program doesn't change School Test Scores" and my lead paragraph would introduce the claim made by the superintendent and bring forth simple, non-technical data to show how the mean scores don't change after the implementation of the reading program. There is little to no evidence to support the superintendent's claim after looking at the data.


### Reflection Question 6:
What metrics or outcomes beyond test scores might be important to track for assessing reading performance?

Besides test scores, I think it could be valuable to look at metrics that can also impact test scores such as the average physical activity the students get per day, average homework the students complete, and metrics such as retention rates.
